CF Registry Analysis

### 2024-05-07 Christine Chan/CF Registry/cf_registry_survival_analysis.qmd
survival analysis mostly finalized

# CFRD screening rates

@@

## Incidence and prevalence

### Full cohort

### By age group

## Screening rates

@@

```{r}
annual_split <- survSplit(
  Surv(as.numeric(tstart), as.numeric(tstop), as.numeric(CFRD) == 2,
    origin = as.numeric(og)
  ) ~ .,
  data = annual, cut = c(2008, 2013, 2018), episode = "tgroup"
)
annual_split$tgroup <- factor(annual_split$tgroup,
  levels = 1:3, labels = c("[2003,2009)", "[2009,2014)", "[2014,2019)")
)
y <- "Surv(as.numeric(tstart), as.numeric(tstop), event)"
```

@@

```{r results='asis'}
invisible(lapply(time_varying, function(v) {
  # Formulas
  new_f <- as.formula(paste0(
    y, "~", paste0(xvars[-which(xvars == v)], collapse = "+"),
    paste0("+", v, "*strata(tgroup)")
  ))
  new_f_int <- as.formula(paste0(
    y, "~", paste0(xvars[-which(xvars == v)], collapse = "+"),
    paste0("+", v, "*tgroup")
  ))
  new_f_emm <- as.formula(paste0(
    "~", paste0(v, "|tgroup")
  ))
  # This could be done more efficiently but it seems to work
  # Fit one stratified model and one with a regular interaction term
  cox_res_split <- coxph(new_f, data = annual_split, cluster = eDWID)
  cox_res_split_int <- coxph(new_f_int, data = annual_split, cluster = eDWID)
  # Use the regular interaction term to get means for each time period
  cat("\n")
  cat("####", v)
  cat("\n")
  cat("\n")
  cox_res_split %>%
    tbl_regression() %>%
    add_n(location = "level") %>%
    add_nevent(location = "level") %>%
    print()
  cat("\n")
  cat("\n")
  emm <- emmeans(cox_res_split_int, new_f_emm,
    nuisance = xvars[-which(xvars == v)]
  )
  emm <- data.frame(emm)
  # Plot base model (without time varying)
  ggcoef_model(cox_res_split,
    show_p_values = F, add_reference_rows = F, signif_stars = F, colour = F,
    include = xvars[-which(xvars == v)]
  ) %>%
    print()
  cat("\n")
  cat("\n")
  # Print EMMs
  print(kable(emm, digits = 3))
  cat("\n")
  cat("\n")
  # Plot marginal means
  time_varying_forest_plot(emm, x = "emmean", var = v, var_label = labels[v]) %>%
    print()
  cat("\n")
  cat("\n")
}))
```

>>>>

# Incidence

## Full cohort

## By age group

# Screening rates

## By age group

## Who is getting OGTTs?

```{r}
# How many OGTTs are non-CFRD people getting 2016-2018
ogtts <- annual %>%
  filter(ReviewYear %in% 2016:2018, CFRD == "CFRD-") %>%
  group_by(eDWID) %>%
  summarise(
    `Age Group` = last(na.omit(age_group)),
    `OGTTs` = sum(!is.na(A_twohour_bloodglucose) | !is.na(A_ogttfast_bloodglucose))
  )
# Add demographics
ogtts$eDWID <- as.numeric(ogtts$eDWID)
ogtts <- left_join(ogtts, demogs, by = join_by(eDWID))
label(ogtts$eDWID) <- "ID"
# Format OGTT counts
ogtts$OGTTs <- factor(ogtts$OGTTs,
  levels = 0:3, labels = c(
    "No OGTT", "At least 1 OGTT", "At least 1 OGTT", "At least 1 OGTT"
  )
)
# Compare no OGTT to at least 1 OGTT
ogtts %>% 
  select(-eDWID) %>%
  tbl_summary(by=OGTTs,missing_text = "Missing") %>%
  add_p()
```

@@

@@

```{r}
annual$tstart <- as.numeric(annual$tstart)
annual$tstop <- as.numeric(annual$tstop)
annual_split <- survSplit(
  Surv(as.numeric(tstart), as.numeric(tstop), as.numeric(CFRD) == 2,
    origin = as.numeric(og)
  ) ~ .,
  data = annual,
  cut = c(
    as.numeric(ymd("2008-12-31")) - as.numeric(og),
    as.numeric(ymd("2013-12-31")) - as.numeric(og)
  ),
  episode = "tgroup"
)
annual_split$tgroup <- factor(annual_split$tgroup,
  levels = 1:3, labels = c("[2003,2009)", "[2009,2014)", "[2014,2019)")
)
y <- "Surv(tstart, tstop, event)"
```

@@

```{r results='asis'}
invisible(lapply(time_varying, function(v) {
  # Formulas
  new_f <- as.formula(paste0(
    y, "~", paste0(xvars[-which(xvars == v)], collapse = "+"),
    paste0("+", v, "*strata(tgroup)")
  ))
  new_f_emm <- as.formula(paste0(
    "~", paste0(v, "|tgroup")
  ))
  # This could be done more efficiently but it seems to work
  # Fit one stratified model and one with a regular interaction term
  cox_res_split <- coxph(new_f, data = annual_split, cluster = eDWID)
  # Use the regular interaction term to get means for each time period
  cat("\n")
  cat("####", v)
  cat("\n")
  cat("\n")
  cox_res_split %>%
    tbl_regression() %>%
    add_n(location = "level") %>%
    add_nevent(location = "level") %>%
    print()
  cat("\n")
  cat("\n")
  emm <- emmeans(cox_res_split, new_f_emm, nuisance = xvars[-which(xvars == v)])
  emm <- data.frame(emm)
  # Plot base model (without time varying)
  ggcoef_model(cox_res_split,
    show_p_values = F, add_reference_rows = F, signif_stars = F, colour = F,
    include = xvars[-which(xvars == v)]
  ) %>%
    print()
  cat("\n")
  cat("\n")
  # Print EMMs
  print(kable(emm, digits = 3))
  cat("\n")
  cat("\n")
  # Plot marginal means
  time_varying_forest_plot(emm, x = "emmean", var = v, var_label = labels[v]) %>%
    print()
  cat("\n")
  cat("\n")
}))
```
